Meta analysis of differential expression results from AMP-AD (Control vs AD, reprocessed RNASeq)

### 2017-03-14 code/Rmd/MetaAnalysis.Rmd
temp change

tmp1 = rbindlist(diff.exp, use.names = T, fill = T) %>%
  dplyr::filter(gene.id %in% Reduce(intersect, diff.exp.gl)) %>%
  dplyr::select(-hgnc_symbol) %>%
  left_join(Ensemble2HGNC)
tmp1$hgnc_symbol[tmp1$ensembl_gene_id == 'ENSG00000275830'] = 'LOC101927646'

tmp = group_by(tmp1, gene.id) %>%
  summarise(count = sum(logFC/abs(logFC))) %>%
  filter(abs(count) == 7) %>%
  dplyr::mutate(GeneID = gene.id) %>%
  tidyr::separate(gene.id, c('ensembl_gene_id','position'), sep = '\\.') %>%
  left_join(Ensemble2HGNC)
writeLines(paste('Differentially expressed genes in all 7 brain regions at an FDR of 0.05 and fold change of 1.2 are',
                 length(unique(tmp$GeneID))))
kable(data.frame(ngenes = length(unique(tmp$GeneID)),
                 hgnc_symbol = tmp$hgnc_symbol %>% unique %>% paste(collapse = ',')))

writeLines('Top six differentially expressed genes are')
tmp.eff.sz = group_by(tmp1, gene.id) %>% 
  summarise(es = sum(logFC, na.rm = T)) %>% 
  top_n(6, abs(es))

pl = list()
p = ggplot(tmp1 %>% 
             filter(gene.id %in% tmp.eff.sz$gene.id[tmp.eff.sz$es>0]), 
           aes(x = Region, y = logFC, fill = Study)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top')
pl[[1]] = p

p = ggplot(tmp1 %>% filter(gene.id %in% tmp.eff.sz$gene.id[tmp.eff.sz$es<0]), aes(x = Region, y = logFC, fill = Study)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top')
pl[[2]] = p

@@

### Pathway enrichment analysis
```{r pathway.enrich}
setsToTest = diffexp %>%
  filter(adj.P.Val <= 0.05) %>%
  dlply(.(Tissue, Direction), .fun = function(x){
    unique(x$ensembl_gene_id)
  })

@@

# Get Background genes
backgroundGenes = Ensemble2HGNC$hgnc_symbol[Ensemble2HGNC$ensembl_gene_id %in% rownames(expr)] %>% unique()

>>>>

tmp1 = rbindlist(diff.exp, use.names = T, fill = T) %>%
  dplyr::filter(gene.id %in% Reduce(c, diff.exp.gl)) %>%
  dplyr::select(-hgnc_symbol) %>%
  left_join(Ensemble2HGNC)
tmp1$hgnc_symbol[tmp1$ensembl_gene_id == 'ENSG00000275830'] = 'LOC101927646'

tmp1$Direction = 'NONE'
tmp1$Direction[tmp1$adj.P.Val <= 0.05 & tmp1$logFC >= log2(1.2)] = 'UP'
tmp1$Direction[tmp1$adj.P.Val <= 0.05 & tmp1$logFC <= -log2(1.2)] = 'DOWN'

tmp = tmp1 %>%
  dplyr::filter(hgnc_symbol != '') %>%
  group_by(hgnc_symbol) %>%
  summarise(count = sum(Direction != 'NONE')) %>%
  filter(count >= 5) 
writeLines(paste('Differentially expressed genes in any 5 brain regions at an FDR of 0.05 and fold change of 1.2 are',
                 length(unique(tmp$hgnc_symbol))))
kable(data.frame(ngenes = length(unique(tmp$hgnc_symbol)),
                 hgnc_symbol = tmp$hgnc_symbol %>% unique %>% paste(collapse = ',')))

writeLines('Top six differentially expressed genes are')
tmp.eff.sz = filter(tmp1, hgnc_symbol %in% tmp$hgnc_symbol) %>%
  group_by(hgnc_symbol) %>% 
  summarise(es = sum(logFC, na.rm = T)) %>% 
  top_n(6, abs(es))

pl = list()
p = ggplot(tmp1 %>% 
             filter(hgnc_symbol %in% tmp.eff.sz$hgnc_symbol[tmp.eff.sz$es>2.6]), 
           aes(x = Region, y = logFC, color = Study, fill = Direction)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top') + scale_fill_manual(values = c('grey', 'red'))
pl[[1]] = p

p = ggplot(tmp1 %>% 
             filter(hgnc_symbol %in% tmp.eff.sz$hgnc_symbol[tmp.eff.sz$es<2.6]), 
           aes(x = Region, y = logFC, color = Study, fill = Direction)) + geom_bar(stat = 'identity')
p = p + facet_grid(hgnc_symbol~.) + geom_errorbar(aes(ymin=CI.L, ymax=CI.R), width = 0.2)
p = p + theme(legend.position = 'top')  + scale_fill_manual(values = c('green','grey','red'))
pl[[2]] = p

@@

### Pathway enrichment analysis
```{r pathway.enrich}
setsToTest = rbindlist(diff.exp, use.names = T, fill = T) %>%
  filter(adj.P.Val <= 0.05, Direction != 'NONE') %>%
  dlply(.(Region, Direction), .fun = function(x){
    unique(x$hgnc_symbol)
  })

@@

# Get Background genes
backgroundGenes = unique(rbindlist(diff.exp, use.names = T, fill = T) %>% 
                           dplyr::select(hgnc_symbol) %>%
                           unlist)
